### 2018-10-15 Manuscript/Manuscript.tex

The original extent LIP polygons were assigned a plate ID corresponding to a tectonic unit on Earth using the polygons of \citet{Torsvik2016a} for the Phanerozoic. The LIP polygons and tectonic units were reconstructed from 520~Ma to the present utilizing the paleogeographic model of \citet{Torsvik2016a} in the spin axis reference frame (anchor plate ID of 1). This paleogeographic model was updated to include a revision to Ordovician Laurentia \citep{Swanson-Hysell2017a} and the Paleozoic of Asia \citep{Domeier2018a}. Reconstructions and area calculations within latitude bands utilized the pyGPlates function library and custom Python scripts documented within a Jupyter notebook that reproduces the analysis and development of the associated visualizations. The total LIP area and the LIP area reconstructed within the tropical rain belt (15\textdegree~S to 15\textdegree~N) were calculated for the various decay and burial scenarios at a resolution of 5~Myr (Fig. \ref{fig:LIP_area}).

To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic Eon (modified from \citealp{Royer2004a}; Fig. \ref{fig:LIP_area}). The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state. The latitudinal extent of land ice is reflective of Earth's overall climate and allows for the delineation of glacial and non-glacial climate states. The land ice record is an imperfect tracke, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals. Nevertheless, it forms a physical record of Earth's climate state through time (Fig. \ref{fig:LIP_area}). We take two approaches for comparison between the LIP area reconstruction and the record of ice extent (Fig. \ref{fig:LIP_area}). The first to to look at the correlation between LIP area and the extent of ice away from the pole (Fig. \ref{fig:LIP_correlation}). The second is to consider the overlap between intervals of high LIP area in the tropics (defined as LIP area in the tropics \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree $\;$from the poles). This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.

@@

OLD TEXT FROM THE SCIENCE MANUSCRIPT THAT NEEDS TO BE UPDATED IN ACCORDANCE WITH NEW RESULTS
In the “decay” scenario, we observe two primary peaks and one minor peak in calculated LIP area within the tropics (Fig. S4B). The first primary peak starting ca. 377 Ma is associated with the emplacement of the Kola-Dnieper and the second primary peak starting ca. 251 Ma is associated with the emplacement of the Central Atlantic magmatic province (CAMP). A Cenozoic peak is associated with both the ca. 30 Ma emplacement of the Afro-Arabian (Afar) LIP as well as the effect of the earlier drift of the Deccan LIP into the tropics. In the “burial” scenario, the peak associated with the CAMP is transient, but the other peaks remain (Fig. S4).

I THINK HERE IN THE RESULTS IS WHERE THE MONTE CARLO APPROACH CAN BE ELUCIDATED. WE CAN SUCCINCTLY MAKE THE POINT THAT DIRECT CROSS-CORRELATION ISN'T APPROPRIATE WITHOUT SUCH NULL HYPOTHESIS TESTING. WE CAN PRESENT THAT THERE IS A WEEK POSITIVE CORRELATION WITH d2+b2 AND THEN DESCRIBE HOW THE RANDOM GENERATE GLACIAL INTERVAL APPROACH SHOWS THAT IT IS NOT SIGNIFICANT. THIS TEXT FROM THE FIG 4 CAPTION IS A START:
\textit{The only scenario for which there is a positive correlation coefficient related to ice extent and LIP area is that of a relatively slow decay rate (d2) and complete burial of LIPs associated with rifting (b2). This positive correlation is weak (\textit{r}=0.10) and is not significant as randomly timed glacial intervals correlate better than the record 28$\%$ of the time. With an associated p-value of 0.28, the null hypothesis that glacial intervals not related to LIP area in the tropics cannot be rejected.}

@@

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
	\caption{\textbf{A)} Paleolatitude of emplacement (centroid) and the initial area estimate of each LIP in the compilation. \textbf{B)} Total LIP area through time for the different post-emplacement models. \textbf{C)} Tropical LIP area through time for the different post-emplacement models. The arrows indicate tropical LIP area for the ca. 720~Ma Franklin LIP and the ca. 1109~Ma Umkondo LIP. \textbf{D)} Heat map for the preferred post-emplacement model showing the latitudinal distribution of LIP area. \textbf{E)} Latitudinal extent of land ice away from the poles through time.}
	\label{fig:LIP_area}
\end{center}
\end{figure}

>>>>

The original extent LIP polygons were assigned a plate ID corresponding to a tectonic unit on Earth using the polygons of \citet{Torsvik2016a} for the Phanerozoic. The LIP polygons and tectonic units were reconstructed from 520~Ma to the present utilizing the paleogeographic model of \citet{Torsvik2016a} in the spin axis reference frame (anchor plate ID of 1). This paleogeographic model was updated to include a revision to Ordovician Laurentia \citep{Swanson-Hysell2017a} and the Paleozoic of Asia \citep{Domeier2018a}. Reconstructions and area calculations within latitude bands utilized the pyGPlates function library and custom Python scripts documented within a Jupyter notebook that reproduces the analysis and development of the associated visualizations. The total LIP area and the LIP area reconstructed within the tropical rain belt (15\textdegree~S to 15\textdegree~N) were calculated for the various decay and burial scenarios at a resolution of 5~Myr (Fig. \ref{fig:LIP_area}B and C).

To evaluate the relationship between Earth's climate state and tropical LIP area, we compared these areas to a compilation of the latitudinal extent of continental ice sheets through time over the Phanerozoic Eon (modified from \citealp{Royer2004a}; Fig. \ref{fig:LIP_area}E). The goal in doing so is to evaluate the hypothesis that there is a correlation between LIP area in the tropics and Earth's long-term climate state. The latitudinal extent of land ice is reflective of Earth's overall climate and allows for the delineation of glacial and non-glacial climate states. The land ice record is an imperfect tracker, as it is insensitive to changes in temperatures during non-glacial intervals and is influenced by additional factors such as the physical geography of the continents during glacial intervals. Nevertheless, it forms a physical record of Earth's climate state through time. We take two approaches for comparison between the LIP area reconstruction and the record of ice extent. The first is to look at the correlation between LIP area and the extent of ice away from the pole (Fig. \ref{fig:LIP_correlation}). The second is to consider the overlap between intervals of high LIP area in the tropics (defined as LIP area in the tropics \textgreater30\% of the maximum in a given post-emplacement model) and intervals of glacial climate (defined as ice extent \textgreater10\textdegree\xspace from the poles). This overlap approach places less emphasis on the specific magnitudes of the compiled ice extent and LIP area records.

@@

In the `decay 1' scenario, we observe three primary peaks and one minor peak in the calculated LIP area within the tropics (Fig. \ref{fig:LIP_area}C). The first primary peak ca. 510~Ma is associated with the emplacement of the Kalkarindji LIP, the second primary peak ca. 380~Ma is associated with the emplacement of the Kola-Dnieper LIP, the third primary peak ca. 200~Ma is associated with the emplacement of the CAMP, and the minor Cenozoic peak is associated with both the ca. 30~Ma emplacement of the Afar LIP as well as the earlier drift of the ca. 66~Ma Deccan LIP into the tropics (Fig. \ref{fig:LIP_area}A). When we account for burial (`decay 1 + burial 1' and `decay 1 + burial 2' scenarios), only the latter two of these four peaks are affected -- the ca. 200~Ma peak is attenuated/removed due to the partial/complete removal of the CAMP, and the Cenozoic peak is attenuated due to the partial/complete removal of the Afar LIP. However, after accounting for burial, a minor area of LIPs remain in the tropics from ca. 130~Ma onwards, due to the EQUAMP, Caribbean-Colombian, and Deccan LIPs. Using the longer decay half-life of 120~Myr (the `decay 2 + burial 2' scenario) increases the area of LIPs in the tropics at any given time step, and has the effect of extending the duration of each peak.

The only scenario which results in a positive Pearson correlation coefficient (0.10) between LIP area in the tropics (Fig. \ref{fig:LIP_area}C) and the ice extent record (Fig. \ref{fig:LIP_area}E) is the scenario with the slow decay rate and complete burial of LIPs associated with rifting (the `decay 2 + burial 2' scenario; Fig. \ref{fig:LIP_correlation}). All other scenarios yield a weak negative correlation coefficient. Qualitatively, the weak positive correlation of the `decay 2 + burial 2' scenario relative to the other scenarios can be primarily attributed to the complete removal of the CAMP, which was emplaced during an extended interval of ice-free conditions, as well as the effect of the longer decay half-life extending the duration of the earlier two peaks, such that they overlap more with the Late Ordivician and Permo-Carboniferous glacial intervals.

To assess the statistical significance of the raw Pearson correlation coefficients,

I THINK HERE IN THE RESULTS IS WHERE THE MONTE CARLO APPROACH CAN BE ELUCIDATED. WE CAN SUCCINCTLY MAKE THE POINT THAT DIRECT CROSS-CORRELATION ISN'T APPROPRIATE WITHOUT SUCH NULL HYPOTHESIS TESTING. WE CAN PRESENT THAT THERE IS A WEAK POSITIVE CORRELATION WITH d2+b2 AND THEN DESCRIBE HOW THE RANDOM GENERATE GLACIAL INTERVAL APPROACH SHOWS THAT IT IS NOT SIGNIFICANT. THIS TEXT FROM THE FIG 4 CAPTION IS A START:
\textit{The only scenario for which there is a positive correlation coefficient related to ice extent and LIP area is that of a relatively slow decay rate (d2) and complete burial of LIPs associated with rifting (b2). This positive correlation is weak (\textit{r}=0.10) and is not significant as randomly timed glacial intervals correlate better than the record 28$\%$ of the time. With an associated p-value of 0.28, the null hypothesis that glacial intervals not related to LIP area in the tropics cannot be rejected.}

@@

\begin{figure}[h!]
\begin{center}
	\includegraphics[width=0.9\textwidth]{Manuscript/Figures/LIP_Areas.pdf}
	\caption{\textbf{A)} Paleolatitude of emplacement (centroid) and the initial area estimate of each LIP in the compilation. Car.-Col. = Caribbean-Colombian. \textbf{B)} Total LIP area through time for the different post-emplacement models. Only the `no decay' model excludes pre-Phanerozoic LIPs. \textbf{C)} Tropical LIP area through time for the different post-emplacement models. The arrows indicate tropical LIP area for the ca. 720~Ma Franklin LIP and the ca. 1109~Ma Umkondo LIP. \textbf{D)} Heat map for the preferred post-emplacement model showing the latitudinal distribution of LIP area. \textbf{E)} Latitudinal extent of land ice away from the poles through time.}
	\label{fig:LIP_area}
\end{center}
\end{figure}
